Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriterial problems

### 2024-06-17 2024/OPTIMA/OPTIMA_2024_MCO.tex
add titlerunning
fix references list

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
\title{Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriterial problems}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Sergey Konnov\orcidID{0009-0003-4590-0870} \and
Evgeny Kozinov \Letter\orcidID{0000-0001-6776-0096} \and
Konstantin Barkalov \Letter\orcidID{0000-0001-5273-2471} \and
Vladimir Grishagin\orcidID{0000-0002-2884-3670}}
%
\authorrunning{S. Konnov, E. Kozinov, K. Barkalov and V. Grishagin}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.

@@

%
\begin{abstract}
To solve a multicriterial optimization problem, it is necessary to find a whole set of parameters values corresponding to non-dominated criteria values (Pa-reto set). The complexity of these problems increases significantly in the case of the criteria are multiextremal. In the paper discusses several varia-tions of the multicriteria optimization algorithm using the black box method with multiextremal criteria. The developed algorithms are based on an infor-mation-statistical approach and a machine learning procedure used to in-crease the efficiency of constructing the Pareto set. As a novelty, the authors proposed a new approach to using the constructed machine learning models. The method involves using the calculated probability of a trial points belong-ing to the Pareto set to assess the prospects one. The effectiveness of differ-ent variations of the developed algorithms estimated in a representative com-putational experiment.

>>>>

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
\title{Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriterial problems\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no. FSWR-2023-0034, and by by the Research and Education Mathematics Centre ``Mathematics for Future Technologies'', project no. 075-02-2024-1439.} }
%
\titlerunning{Algorithms for solving multicriterial problems}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Sergey Konnov\orcidID{0009-0003-4590-0870} \and
Evgeny Kozinov \Letter\orcidID{0000-0001-6776-0096} \and
Konstantin Barkalov \Letter\orcidID{0000-0001-5273-2471} \and
Vladimir Grishagin\orcidID{0000-0002-2884-3670}}
%
\authorrunning{S. Konnov, E. Kozinov, K. Barkalov and V. Grishagin}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.

@@

%
\begin{abstract}
To solve a multicriterial optimization problem, it is necessary to find a whole set of parameters values corresponding to non-dominated criteria values (Pareto set). The complexity of these problems increases significantly in the case of the criteria are multiextremal. In the paper discusses several varia-tions of the multicriteria optimization algorithm using the black box method with multiextremal criteria. The developed algorithms are based on an information-statistical approach and a machine learning procedure used to increase the efficiency of constructing the Pareto set. As a novelty, the authors proposed a new approach to using the constructed machine learning models. The method involves using the calculated probability of a trial points belonging to the Pareto set to assess the prospects one. The effectiveness of different variations of the developed algorithms estimated in a representative computational experiment.
